<!-- combined two files, so there are duplicates.  RCLM-presentation.Rmd has more efficient (compact) code. -->
<!-- source data is from csv files, not the "original" SPSS sav files -->

Summary of Fall 2013 survey data
========================================================

In this script we import the Fall 2013 MASC 351 survey data and calculate simple descriptive statistics.

First we import the data from reduced csv files that were exported from Qualtrics through SPSS format intermediary.

```{r echo=FALSE}
rm(list=ls())
setwd("~/Dropbox/KATE Admin/KATE-Fall-2013/")
library("likert")
diversity.1 <- read.csv("Diversity_Part_1_Fall_2013_reduced.csv", header=TRUE, sep=",", quote="\"")
diversity.2 <- read.csv("Diversity_Part_2_Fall_2013_reduced.csv", header=TRUE, sep=",", quote="\"")
str(diversity.1)
str(diversity.2)
```

Then we process the data from the first survey

```{r echo=FALSE}
part1.Q7 <- data.frame(diversity.1$Q7_1, diversity.1$Q7_2, diversity.1$Q7_3, diversity.1$Q7_4, diversity.1$Q7_5, diversity.1$Q7_6, diversity.1$Q7_7)
names(part1.Q7) <- c("Q7_1","Q7_2","Q7_3","Q7_4","Q7_5","Q7_6","Q7_7")

part1.Q4 <- data.frame(diversity.1$Q4_1[1:26], diversity.1$Q4_2[1:26], diversity.1$Q4_3[1:26], diversity.1$Q4_4[1:26], diversity.1$Q4_5[1:26], diversity.1$Q4_6[1:26], diversity.1$Q4_7[1:26])
names(part1.Q4) <- c("Q4_1","Q4_2","Q4_3","Q4_4","Q4_5","Q4_6","Q4_7")
```

Followed by the data from the second survey

```{r echo=FALSE}
part2.Q2 <- data.frame(diversity.2$Q2_1,
                       diversity.2$Q2_2,
                       diversity.2$Q2_3,
                       diversity.2$Q2_4,
                       diversity.2$Q2_5,
                       diversity.2$Q2_6,
                       diversity.2$Q2_7,
                       diversity.2$Q2_8,
                       diversity.2$Q2_9,
                       diversity.2$Q2_10,
                       diversity.2$Q2_11,
                       diversity.2$Q2_12,
                       diversity.2$Q2_13,
                       diversity.2$Q2_14)
names(part2.Q2) <- c("Q2_1","Q2_2","Q2_3","Q2_4","Q2_5","Q2_6","Q2_7","Q2_8","Q2_9","Q2_10","Q2_11","Q2_12","Q2_13","Q2_14")

part2.Q4 <- data.frame(diversity.2$Q4_1[1:24],
                       diversity.2$Q4_2[1:24],
                       diversity.2$Q4_3[1:24],
                       diversity.2$Q4_4[1:24],
                       diversity.2$Q4_5[1:24],
                       diversity.2$Q4_6[1:24],
                       diversity.2$Q4_7[1:24],
                       diversity.2$Q4_8[1:24],
                       diversity.2$Q4_9[1:24],
                       diversity.2$Q4_10[1:24],
                       diversity.2$Q4_11[1:24],
                       diversity.2$Q4_12[1:24],
                       diversity.2$Q4_13[1:24],
                       diversity.2$Q4_14[1:24])
names(part2.Q4) <- c("Q4_1","Q4_2","Q4_3","Q4_4","Q4_5","Q4_6","Q4_7","Q4_8","Q4_9","Q4_10","Q4_11","Q4_12","Q4_13","Q4_14")
```

The values of the survey are based on a Likert scale, 1 to 4 with the following values:

"No change" (1), "made me rethink my awareness of diversity issues" (2), "improved my awareness of diversity issues somewhat" (3), "improved my awareness of diversity issues a lot" (4)


## We replace the integer values with actual responses ##

When you export the results from Qualtrics, they are given as integer values.  Thus, first we change the values from integers to characters.

```{r echo=FALSE}
part1.Q4.char <- sapply(part1.Q4, as.character)
part1.Q7.char <- sapply(part1.Q7, as.character)
part2.Q2.char <- sapply(part2.Q2, as.character)
part2.Q4.char <- sapply(part2.Q4, as.character)
```

The we replace the numbers with the appropriate character strings.
```{r echo=FALSE}
part1.Q7.char.temp <- gsub("1","No change",part1.Q7.char)
part1.Q7.char.temp <- gsub("2","Made me rethink",part1.Q7.char.temp)
part1.Q7.char.temp <- gsub("3","Improved somewhat",part1.Q7.char.temp)
part1.Q7.char.temp <- gsub("4","Improved a lot",part1.Q7.char.temp)
part1.Q7.char <- data.frame(part1.Q7.char.temp)
rm(part1.Q7.char.temp)

part1.Q4.char.temp <- gsub("1","No change",part1.Q4.char)
part1.Q4.char.temp <- gsub("2","Made me rethink",part1.Q4.char.temp)
part1.Q4.char.temp <- gsub("3","Improved somewhat",part1.Q4.char.temp)
part1.Q4.char.temp <- gsub("4","Improved a lot",part1.Q4.char.temp)
part1.Q4.char <- data.frame(part1.Q4.char.temp)
rm(part1.Q4.char.temp)

part2.Q2.char.temp <- gsub("1","No change",part2.Q2.char)
part2.Q2.char.temp <- gsub("2","Made me rethink",part2.Q2.char.temp)
part2.Q2.char.temp <- gsub("3","Improved somewhat",part2.Q2.char.temp)
part2.Q2.char.temp <- gsub("4","Improved a lot",part2.Q2.char.temp)
part2.Q2.char <- data.frame(part2.Q2.char.temp)
rm(part2.Q2.char.temp)

part2.Q4.char.temp <- gsub("1","No change",part2.Q4.char)
part2.Q4.char.temp <- gsub("2","Made me rethink",part2.Q4.char.temp)
part2.Q4.char.temp <- gsub("3","Improved somewhat",part2.Q4.char.temp)
part2.Q4.char.temp <- gsub("4","Improved a lot",part2.Q4.char.temp)
part2.Q4.char <- data.frame(part2.Q4.char.temp)
rm(part2.Q4.char.temp)
```

All dataframes need the same levels to convert them to likert objects
```{r echo=FALSE}
part1.Q4.char$Q4_1 <- ordered(part1.Q4$Q4_1, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part1.Q4.char$Q4_2 <- ordered(part1.Q4$Q4_2, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part1.Q4.char$Q4_3 <- ordered(part1.Q4$Q4_3, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part1.Q4.char$Q4_4 <- ordered(part1.Q4$Q4_4, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part1.Q4.char$Q4_5 <- ordered(part1.Q4$Q4_5, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part1.Q4.char$Q4_6 <- ordered(part1.Q4$Q4_6, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part1.Q4.char$Q4_7 <- ordered(part1.Q4$Q4_7, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))

part1.Q7.char$Q7_1 <- ordered(part1.Q7$Q7_1, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part1.Q7.char$Q7_2 <- ordered(part1.Q7$Q7_2, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part1.Q7.char$Q7_3 <- ordered(part1.Q7$Q7_3, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part1.Q7.char$Q7_4 <- ordered(part1.Q7$Q7_4, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part1.Q7.char$Q7_5 <- ordered(part1.Q7$Q7_5, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part1.Q7.char$Q7_6 <- ordered(part1.Q7$Q7_6, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part1.Q7.char$Q7_7 <- ordered(part1.Q7$Q7_7, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))

part2.Q2.char$Q2_1  <- ordered(part2.Q2$Q2_1, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_2  <- ordered(part2.Q2$Q2_2, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_3  <- ordered(part2.Q2$Q2_3, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_4  <- ordered(part2.Q2$Q2_4, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_5  <- ordered(part2.Q2$Q2_5, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_6  <- ordered(part2.Q2$Q2_6, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_7  <- ordered(part2.Q2$Q2_7, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_8  <- ordered(part2.Q2$Q2_8, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_9  <- ordered(part2.Q2$Q2_9, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_10 <- ordered(part2.Q2$Q2_10, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_11 <- ordered(part2.Q2$Q2_11, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_12 <- ordered(part2.Q2$Q2_12, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_13 <- ordered(part2.Q2$Q2_13, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_14 <- ordered(part2.Q2$Q2_14, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))

part2.Q4.char$Q4_1  <- ordered(part2.Q4$Q4_1, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_2  <- ordered(part2.Q4$Q4_2, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_3  <- ordered(part2.Q4$Q4_3, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_4  <- ordered(part2.Q4$Q4_4, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_5  <- ordered(part2.Q4$Q4_5, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_6  <- ordered(part2.Q4$Q4_6, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_7  <- ordered(part2.Q4$Q4_7, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_8  <- ordered(part2.Q4$Q4_8, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_9  <- ordered(part2.Q4$Q4_9, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_10 <- ordered(part2.Q4$Q4_10, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_11 <- ordered(part2.Q4$Q4_11, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_12 <- ordered(part2.Q4$Q4_12, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_13 <- ordered(part2.Q4$Q4_13, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_14 <- ordered(part2.Q4$Q4_14, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
```

Assign the question to each column
```{r echo=FALSE}
names(part1.Q4.char) <- c("Readings and questions from the Polya textbook",
                          "Readings and questions from the Responding to Diversity textbook",
                          "Presentations and discussions in class about culture and diversity",
                          "Dinner Problem Activity: a. Solving the problem",
                          "Dinner Problem Activity: b. Responding to student misconceptions",
                          "Dinner Problem Activity: c. Planning a lesson",
                          "Dinner Problem Activity: d. Responding to student questions")

names(part1.Q7.char) <- c("Readings and questions from the Polya textbook",
                          "Readings and questions from the Responding to Diversity textbook",
                          "Presentations and discussions in class about culture and diversity",
                          "Dinner Problem Activity: a. Solving the problem",
                          "Dinner Problem Activity: b. Responding to student misconceptions",
                          "Dinner Problem Activity: c. Planning a lesson",
                          "Dinner Problem Actitivy: d. Responding to student questions")

names(part2.Q2.char) <- c("Diversity Readings (Ladson-Billings; Milner)",
                          "Guest Speaker: Dr. Chance Lewis",
                          "Soda Can Problem: a. Solving the problem",
                          "Soda Can Problem: b. Responding to student misconceptions",
                          "Soda Can Problem: c. Planning a lesson",
                          "Soda Can Problem: d. Responding to student",
                          "Meet the Middle Grade Students in SL",
                          "Tutoring the Middle Grade Students in SL",
                          "Analysis of Language Moves from your tutoring",
                          "Guest Speaker: Dr. Kathryn McKenzie",
                          "Food Drive Problem: a. Solving the problem",
                          "Food Drive Problem: b. Responding to student misconceptions",
                          "Food Drive Problem: c. Planning a lesson",
                          "Food Drive Problem: d. Responding to student questions")

names(part2.Q4.char) <- c("Diversity Readings (Ladson-Billings; Milner)",
                          "Guest Speaker: Dr. Chance Lewis",
                          "Soda Can Problem: a. Solving the problem",
                          "Soda Can Problem: b. Responding to student misconceptions",
                          "Soda Can Problem: c. Planning a lesson",
                          "Soda Can Problem: d. Responding to student",
                          "Meet the Middle Grade Students in SL",
                          "Tutoring the Middle Grade Students in SL",
                          "Analysis of Language Moves from your tutoring",
                          "Guest Speaker: Dr. Kathryn McKenzie",
                          "Food Drive Problem: a. Solving the problem",
                          "Food Drive Problem: b. Responding to student misconceptions",
                          "Food Drive Problem: c. Planning a lesson",
                          "Food Drive Problem: d. Responding to student questions")
```

Convert to Likert objects
```{r echo=FALSE}
part1.Q4.likert <- likert(part1.Q4.char)
part1.Q7.likert <- likert(part1.Q7.char)
part2.Q2.likert <- likert(part2.Q2.char)
part2.Q4.likert <- likert(part2.Q4.char)
```

## Finally, plot the four main questions ##

Part 1, Question 7 "How has each of the following activities CHANGED your AWARENESS about diversity issues?"
```{r echo=FALSE, fig.width=10}
plot(part1.Q7.likert, centered = FALSE, wrap = 25)
```

Part 1, Question 4 "How has each of the following activities CHANGED your KNOWLEDGE ABOUT TEACHING algebra diversity?"
```{r echo=FALSE, fig.width=10}
plot(part1.Q4.likert, centered = FALSE, wrap = 25)
```

Part 2, Question 2 "How has each of the following activities CHANGED your AWARENESS about diversity issues?"
```{r echo=FALSE, fig.width=10}
plot(part2.Q2.likert, centered = FALSE, wrap = 32)
```

Part 2, Question 4 "How has each of the following activities CHANGED your KNOWLEDGE ABOUT TEACHING algebra diversity?"
```{r echo=FALSE, fig.width=10}
plot(part2.Q4.likert, centered = FALSE, wrap = 32)
```

## "Heat" charts ##

Heat charts also provide some summary values based on the 4 integer values as shown previously.

Part 1, Question 7 "How has each of the following activities CHANGED your AWARENESS about diversity issues?"
```{r echo=FALSE, fig.width=10}
plot(part1.Q7.likert, type="heat")
```

Part 1, Question 4 "How has each of the following activities CHANGED your KNOWLEDGE ABOUT TEACHING algebra diversity?"
```{r echo=FALSE, fig.width=10}
plot(part1.Q4.likert, type="heat")
```

Part 2, Question 2 "How has each of the following activities CHANGED your AWARENESS about diversity issues?"
```{r echo=FALSE, fig.width=10}
plot(part2.Q2.likert, type="heat")
```

Part 2, Question 4 "How has each of the following activities CHANGED your KNOWLEDGE ABOUT TEACHING algebra diversity?"
```{r echo=FALSE, fig.width=10}
plot(part2.Q4.likert, type="heat")
```

<!-- ~/Dropbox/KATE Admin/KATE-Fall-2013/Summary-survey-fall-2013.Rmd -->

Summary of Fall 2013 survey data
========================================================

In this script we import the Fall 2013 MASC 351 survey data and calculate simple descriptive statistics.

First we import the data from reduced csv files that were exported from Qualtrics.

```{r echo=FALSE}
rm(list=ls())
setwd("~/Dropbox/KATE Admin/KATE-Fall-2013/")
library("likert")
diversity.1 <- read.csv("Diversity_Part_1_Fall_2013_reduced.csv", header=TRUE, sep=",", quote="\"")
diversity.2 <- read.csv("Diversity_Part_2_Fall_2013_reduced.csv", header=TRUE, sep=",", quote="\"")
str(diversity.1)
str(diversity.2)
```

Then we process the data from the first survey

```{r echo=FALSE}
part1.Q7 <- data.frame(diversity.1$Q7_1, diversity.1$Q7_2, diversity.1$Q7_3, diversity.1$Q7_4, diversity.1$Q7_5, diversity.1$Q7_6, diversity.1$Q7_7)
names(part1.Q7) <- c("Q7_1","Q7_2","Q7_3","Q7_4","Q7_5","Q7_6","Q7_7")

part1.Q4 <- data.frame(diversity.1$Q4_1[1:26], diversity.1$Q4_2[1:26], diversity.1$Q4_3[1:26], diversity.1$Q4_4[1:26], diversity.1$Q4_5[1:26], diversity.1$Q4_6[1:26], diversity.1$Q4_7[1:26])
names(part1.Q4) <- c("Q4_1","Q4_2","Q4_3","Q4_4","Q4_5","Q4_6","Q4_7")
```

Followed by the data from the second survey

```{r echo=FALSE}
part2.Q2 <- data.frame(diversity.2$Q2_1,
                       diversity.2$Q2_2,
                       diversity.2$Q2_3,
                       diversity.2$Q2_4,
                       diversity.2$Q2_5,
                       diversity.2$Q2_6,
                       diversity.2$Q2_7,
                       diversity.2$Q2_8,
                       diversity.2$Q2_9,
                       diversity.2$Q2_10,
                       diversity.2$Q2_11,
                       diversity.2$Q2_12,
                       diversity.2$Q2_13,
                       diversity.2$Q2_14)
names(part2.Q2) <- c("Q2_1","Q2_2","Q2_3","Q2_4","Q2_5","Q2_6","Q2_7","Q2_8","Q2_9","Q2_10","Q2_11","Q2_12","Q2_13","Q2_14")

part2.Q4 <- data.frame(diversity.2$Q4_1[1:24],
                       diversity.2$Q4_2[1:24],
                       diversity.2$Q4_3[1:24],
                       diversity.2$Q4_4[1:24],
                       diversity.2$Q4_5[1:24],
                       diversity.2$Q4_6[1:24],
                       diversity.2$Q4_7[1:24],
                       diversity.2$Q4_8[1:24],
                       diversity.2$Q4_9[1:24],
                       diversity.2$Q4_10[1:24],
                       diversity.2$Q4_11[1:24],
                       diversity.2$Q4_12[1:24],
                       diversity.2$Q4_13[1:24],
                       diversity.2$Q4_14[1:24])
names(part2.Q4) <- c("Q4_1","Q4_2","Q4_3","Q4_4","Q4_5","Q4_6","Q4_7","Q4_8","Q4_9","Q4_10","Q4_11","Q4_12","Q4_13","Q4_14")
```

Finally we calculate some descriptive statistics.  The values are based on a Likert scale, 1 to 4 with the following values:

"No change", "made me rethink my awareness of diversity issues", "improved my awareness of diversity issues somewhat", "improved my awareness of diversity issues a lot"


Part 1, Question 7 "How has each of the following activities CHANGED your AWARENESS about diversity issues?"

Readings and questions from the *Polya* textbook
```{r echo=FALSE}
summary(part1.Q7$Q7_1)
```

Readings and questions from the *Responding to Diversity* textbook
```{r echo=FALSE}
summary(part1.Q7$Q7_2)
```

Presentations and discussions in class about culture and diversity
```{r echo=FALSE}
summary(part1.Q7$Q7_3)
```

Dinner Problem Activity: a. Solving the problem
```{r echo=FALSE}
summary(part1.Q7$Q7_4)
```

Dinner Problem Activity: b. Responding to student misconceptions
```{r echo=FALSE}
summary(part1.Q7$Q7_5)
```

Dinner Problem Activity: c. Planning a lesson
```{r echo=FALSE}
summary(part1.Q7$Q7_6)
```

Dinner Problem Activity: d. Responding to student questions
```{r echo=FALSE}
summary(part1.Q7$Q7_7)
```


Part 1, Question 4 "How has each of the following activities CHANGED your KNOWLEDGE ABOUT TEACHING algebra diversity?"

Readings and questions from the Polya textbook
```{r echo=FALSE}
summary(part1.Q4$Q4_1)
```

Readings and questions from the *Responding to Diversity* textbook
```{r echo=FALSE}
summary(part1.Q4$Q4_2)
```

Presentations and discussions in class about culture and diversity
```{r echo=FALSE}
summary(part1.Q4$Q4_3)
```

Dinner Problem Activity: a. Solving the problem
```{r echo=FALSE}
summary(part1.Q4$Q4_4)
```

Dinner Problem Activity: b. Responding to student misconceptions
```{r echo=FALSE}
summary(part1.Q4$Q4_5)
```

Dinner Problem Activity: c. Planning a lesson
```{r echo=FALSE}
summary(part1.Q4$Q4_6)
```

Dinner Problem Activity: d. Responding to student questions
```{r echo=FALSE}
summary(part1.Q4$Q4_7)
```


Part 2, Question 2 "How has each of the following activities CHANGED your AWARENESS about diversity issues?"

Diversity Readings (*Ladson-Billings; Milner*)
```{r echo=FALSE}
summary(part2.Q2$Q2_1)
```

Guest Speaker: Dr. Chance Lewis
```{r echo=FALSE}
summary(part2.Q2$Q2_2)
```

Soda Can Problem: a. Solving the problem
```{r echo=FALSE}
summary(part2.Q2$Q2_3)
```

Soda Can Problem: b. Responding to student misconceptions
```{r echo=FALSE}
summary(part2.Q2$Q2_4)
```

Soda Can Problem: c. Planning a lesson
```{r echo=FALSE}
summary(part2.Q2$Q2_5)
```

Soda Can Problem: d. Responding to student questions
```{r echo=FALSE}
summary(part2.Q2$Q2_6)
```

Meet the Middle Grade Students in SL
```{r echo=FALSE}
summary(part2.Q2$Q2_7)
```

Tutoring the Middle Grade Students in SL
```{r echo=FALSE}
summary(part2.Q2$Q2_8)
```

Analysis of Language Moves from your tutoring
```{r echo=FALSE}
summary(part2.Q2$Q2_9)
```

Guest Speaker: Dr. Kathryn McKenzie
```{r echo=FALSE}
summary(part2.Q2$Q2_10)
```

Food Drive Problem: a. Solving the problem
```{r echo=FALSE}
summary(part2.Q2$Q2_11)
```

Food Drive Problem: b. Responding to student misconceptions
```{r echo=FALSE}
summary(part2.Q2$Q2_12)
```

Food Drive Problem: c. Planning a lesson
```{r echo=FALSE}
summary(part2.Q2$Q2_13)
```

Food Drive Problem: d. Responding to student questions
```{r echo=FALSE}
summary(part2.Q2$Q2_14)
```


Part 2, Question 4 "How has each of the following activities CHANGED your KNOWLEDGE ABOUT TEACHING algebra diversity?"

Diversity Readings (*Ladson-Billings; Milner*)
```{r echo=FALSE}
summary(part2.Q4$Q4_1)
```

Guest Speaker: Dr. Chance Lewis
```{r echo=FALSE}
summary(part2.Q4$Q4_2)
```

Soda Can Problem: a. Solving the problem
```{r echo=FALSE}
summary(part2.Q4$Q4_3)
```

Soda Can Problem: b. Responding to student misconceptions
```{r echo=FALSE}
summary(part2.Q4$Q4_4)
```

Soda Can Problem: c. Planning a lesson
```{r echo=FALSE}
summary(part2.Q4$Q4_5)
```

Soda Can Problem: d. Responding to student questions
```{r echo=FALSE}
summary(part2.Q4$Q4_6)
```

Meet the Middle Grade Students in SL
```{r echo=FALSE}
summary(part2.Q4$Q4_7)
```

Tutoring the Middle Grade Students in SL
```{r echo=FALSE}
summary(part2.Q4$Q4_8)
```

Analysis of Language Moves from your tutoring
```{r echo=FALSE}
summary(part2.Q4$Q4_9)
```

Guest Speaker: Dr. Kathryn McKenzie
```{r echo=FALSE}
summary(part2.Q4$Q4_10)
```

Food Drive Problem: a. Solving the problem
```{r echo=FALSE}
summary(part2.Q4$Q4_11)
```

Food Drive Problem: b. Responding to student misconceptions
```{r echo=FALSE}
summary(part2.Q4$Q4_12)
```

Food Drive Problem: c. Planning a lesson
```{r echo=FALSE}
summary(part2.Q4$Q4_13)
```

Food Drive Problem: d. Responding to student
```{r echo=FALSE}
summary(part2.Q4$Q4_14)
```


## We replace the integer values with actual responses ##

First we change the values from integers to characters.
```{r echo=FALSE}
part1.Q4.char <- sapply(part1.Q4, as.character)
part1.Q7.char <- sapply(part1.Q7, as.character)
part2.Q2.char <- sapply(part2.Q2, as.character)
part2.Q4.char <- sapply(part2.Q4, as.character)
```

The we replace the numbers with the appropriate character strings.
```{r echo=FALSE}
part1.Q7.char.temp <- gsub("1","No change",part1.Q7.char)
part1.Q7.char.temp <- gsub("2","Made me rethink",part1.Q7.char.temp)
part1.Q7.char.temp <- gsub("3","Improved somewhat",part1.Q7.char.temp)
part1.Q7.char.temp <- gsub("4","Improved a lot",part1.Q7.char.temp)
part1.Q7.char <- data.frame(part1.Q7.char.temp)
rm(part1.Q7.char.temp)

part1.Q4.char.temp <- gsub("1","No change",part1.Q4.char)
part1.Q4.char.temp <- gsub("2","Made me rethink",part1.Q4.char.temp)
part1.Q4.char.temp <- gsub("3","Improved somewhat",part1.Q4.char.temp)
part1.Q4.char.temp <- gsub("4","Improved a lot",part1.Q4.char.temp)
part1.Q4.char <- data.frame(part1.Q4.char.temp)
rm(part1.Q4.char.temp)

part2.Q2.char.temp <- gsub("1","No change",part2.Q2.char)
part2.Q2.char.temp <- gsub("2","Made me rethink",part2.Q2.char.temp)
part2.Q2.char.temp <- gsub("3","Improved somewhat",part2.Q2.char.temp)
part2.Q2.char.temp <- gsub("4","Improved a lot",part2.Q2.char.temp)
part2.Q2.char <- data.frame(part2.Q2.char.temp)
rm(part2.Q2.char.temp)

part2.Q4.char.temp <- gsub("1","No change",part2.Q4.char)
part2.Q4.char.temp <- gsub("2","Made me rethink",part2.Q4.char.temp)
part2.Q4.char.temp <- gsub("3","Improved somewhat",part2.Q4.char.temp)
part2.Q4.char.temp <- gsub("4","Improved a lot",part2.Q4.char.temp)
part2.Q4.char <- data.frame(part2.Q4.char.temp)
rm(part2.Q4.char.temp)
```

All dataframes need the same levels to convert them to likert objects
```{r echo=FALSE}
part1.Q4.char$Q4_1 <- ordered(part1.Q4$Q4_1, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part1.Q4.char$Q4_2 <- ordered(part1.Q4$Q4_2, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part1.Q4.char$Q4_3 <- ordered(part1.Q4$Q4_3, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part1.Q4.char$Q4_4 <- ordered(part1.Q4$Q4_4, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part1.Q4.char$Q4_5 <- ordered(part1.Q4$Q4_5, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part1.Q4.char$Q4_6 <- ordered(part1.Q4$Q4_6, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part1.Q4.char$Q4_7 <- ordered(part1.Q4$Q4_7, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))

part1.Q7.char$Q7_1 <- ordered(part1.Q7$Q7_1, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part1.Q7.char$Q7_2 <- ordered(part1.Q7$Q7_2, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part1.Q7.char$Q7_3 <- ordered(part1.Q7$Q7_3, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part1.Q7.char$Q7_4 <- ordered(part1.Q7$Q7_4, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part1.Q7.char$Q7_5 <- ordered(part1.Q7$Q7_5, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part1.Q7.char$Q7_6 <- ordered(part1.Q7$Q7_6, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part1.Q7.char$Q7_7 <- ordered(part1.Q7$Q7_7, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))

part2.Q2.char$Q2_1  <- ordered(part2.Q2$Q2_1, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_2  <- ordered(part2.Q2$Q2_2, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_3  <- ordered(part2.Q2$Q2_3, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_4  <- ordered(part2.Q2$Q2_4, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_5  <- ordered(part2.Q2$Q2_5, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_6  <- ordered(part2.Q2$Q2_6, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_7  <- ordered(part2.Q2$Q2_7, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_8  <- ordered(part2.Q2$Q2_8, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_9  <- ordered(part2.Q2$Q2_9, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_10 <- ordered(part2.Q2$Q2_10, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_11 <- ordered(part2.Q2$Q2_11, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_12 <- ordered(part2.Q2$Q2_12, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_13 <- ordered(part2.Q2$Q2_13, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q2.char$Q2_14 <- ordered(part2.Q2$Q2_14, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))

part2.Q4.char$Q4_1  <- ordered(part2.Q4$Q4_1, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_2  <- ordered(part2.Q4$Q4_2, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_3  <- ordered(part2.Q4$Q4_3, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_4  <- ordered(part2.Q4$Q4_4, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_5  <- ordered(part2.Q4$Q4_5, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_6  <- ordered(part2.Q4$Q4_6, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_7  <- ordered(part2.Q4$Q4_7, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_8  <- ordered(part2.Q4$Q4_8, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_9  <- ordered(part2.Q4$Q4_9, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_10 <- ordered(part2.Q4$Q4_10, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_11 <- ordered(part2.Q4$Q4_11, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_12 <- ordered(part2.Q4$Q4_12, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_13 <- ordered(part2.Q4$Q4_13, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
part2.Q4.char$Q4_14 <- ordered(part2.Q4$Q4_14, levels=c(1,2,3,4), labels=c("No change","Made me rethink","Improved somewhat","Improved a lot"))
```

Assign the question to each column
```{r echo=FALSE}
names(part1.Q4.char) <- c("Readings and questions from the Polya textbook",
                          "Readings and questions from the Responding to Diversity textbook",
                          "Presentations and discussions in class about culture and diversity",
                          "Dinner Problem Activity: a. Solving the problem",
                          "Dinner Problem Activity: b. Responding to student misconceptions",
                          "Dinner Problem Activity: c. Planning a lesson",
                          "Dinner Problem Activity: d. Responding to student questions")

names(part1.Q7.char) <- c("Readings and questions from the Polya textbook",
                          "Readings and questions from the Responding to Diversity textbook",
                          "Presentations and discussions in class about culture and diversity",
                          "Dinner Problem Activity: a. Solving the problem",
                          "Dinner Problem Activity: b. Responding to student misconceptions",
                          "Dinner Problem Activity: c. Planning a lesson",
                          "Dinner Problem Actitivy: d. Responding to student questions")

names(part2.Q2.char) <- c("Diversity Readings (Ladson-Billings; Milner)",
                          "Guest Speaker: Dr. Chance Lewis",
                          "Soda Can Problem: a. Solving the problem",
                          "Soda Can Problem: b. Responding to student misconceptions",
                          "Soda Can Problem: c. Planning a lesson",
                          "Soda Can Problem: d. Responding to student",
                          "Meet the Middle Grade Students in SL",
                          "Tutoring the Middle Grade Students in SL",
                          "Analysis of Language Moves from your tutoring",
                          "Guest Speaker: Dr. Kathryn McKenzie",
                          "Food Drive Problem: a. Solving the problem",
                          "Food Drive Problem: b. Responding to student misconceptions",
                          "Food Drive Problem: c. Planning a lesson",
                          "Food Drive Problem: d. Responding to student questions")

names(part2.Q4.char) <- c("Diversity Readings (Ladson-Billings; Milner)",
                          "Guest Speaker: Dr. Chance Lewis",
                          "Soda Can Problem: a. Solving the problem",
                          "Soda Can Problem: b. Responding to student misconceptions",
                          "Soda Can Problem: c. Planning a lesson",
                          "Soda Can Problem: d. Responding to student",
                          "Meet the Middle Grade Students in SL",
                          "Tutoring the Middle Grade Students in SL",
                          "Analysis of Language Moves from your tutoring",
                          "Guest Speaker: Dr. Kathryn McKenzie",
                          "Food Drive Problem: a. Solving the problem",
                          "Food Drive Problem: b. Responding to student misconceptions",
                          "Food Drive Problem: c. Planning a lesson",
                          "Food Drive Problem: d. Responding to student questions")
```

Convert to Likert objects
```{r echo=FALSE}
part1.Q4.likert <- likert(part1.Q4.char)
part1.Q7.likert <- likert(part1.Q7.char)
part2.Q2.likert <- likert(part2.Q2.char)
part2.Q4.likert <- likert(part2.Q4.char)
```

## Finally, plot the four main questions ##

Part 1, Question 7 "How has each of the following activities CHANGED your AWARENESS about diversity issues?"
```{r echo=FALSE, fig.width=10}
plot(part1.Q7.likert, centered = FALSE, wrap = 25)
```


Part 1, Question 4 "How has each of the following activities CHANGED your KNOWLEDGE ABOUT TEACHING algebra diversity?"
```{r echo=FALSE, fig.width=10}
plot(part1.Q4.likert, centered = FALSE, wrap = 25)
```

Part 2, Question 2 "How has each of the following activities CHANGED your AWARENESS about diversity issues?"
```{r echo=FALSE, fig.width=10}
plot(part2.Q2.likert, centered = FALSE, wrap = 32)
```

Part 2, Question 4 "How has each of the following activities CHANGED your KNOWLEDGE ABOUT TEACHING algebra diversity?"
```{r echo=FALSE, fig.width=10}
plot(part2.Q4.likert, centered = FALSE, wrap = 32)
```

How about "heat" charts?

Part 1, Question 7 "How has each of the following activities CHANGED your AWARENESS about diversity issues?"
```{r echo=FALSE, fig.width=10}
plot(part1.Q7.likert, type="heat")
```


Part 1, Question 4 "How has each of the following activities CHANGED your KNOWLEDGE ABOUT TEACHING algebra diversity?"
```{r echo=FALSE, fig.width=10}
plot(part1.Q4.likert, type="heat")
```

Part 2, Question 2 "How has each of the following activities CHANGED your AWARENESS about diversity issues?"
```{r echo=FALSE, fig.width=10}
plot(part2.Q2.likert, type="heat")
```

Part 2, Question 4 "How has each of the following activities CHANGED your KNOWLEDGE ABOUT TEACHING algebra diversity?"
```{r echo=FALSE, fig.width=10}
plot(part2.Q4.likert, type="heat")
```
